### 2024-05-22 homework2/paper/paper.tex
organize paper files

\documentclass[runningheads]{paper}

\usepackage{hyperref}   % hyperlinks
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{multirow}
\usepackage{longtable}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

% ------------------------------------------------------------------------------

\begin{document}
%
\title{TODO}
%
\author{Cărămidă Iustina-Andreea - 332CA}
%
\institute{Faculty of Automatic Control and Computer Science \\
University Politehnica of Bucharest \\
\email{iustina.caramida@stud.acs.upb.ro}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    TODO

\keywords{TO \and DO.}
\end{abstract}

% ------------------------------------------------------------------------------
\pagebreak
\section{Introduction}

% ------------------------------------------------------------------------------
\section{Algorithm Design}

% ------------------------------------------------------------------------------

\section{Evaluation}

% ------------------------------------------------------------------------------
\section{Conclusions}

\pagebreak
% ------------------------------------------------------------------------------

% ---- Bibliography ----
\begin{thebibliography}{8}
    \bibitem{}
    \href{https://en.wikipedia.org/wiki/A*_search_algorithm#Complexity}{A* search algorithm - Wikipedia}
    \bibitem{}
    \href{https://en.wikipedia.org/wiki/Hill_climbing}{Hill Climbing - Wikipedia}
    \bibitem{}
    \href{https://curs.upb.ro/2023/course/view.php?id=13749}{Moodle - Artifical Intelligence Course}
    \bibitem{}
    \href{https://www.researchgate.net/profile/Kian-Pokorny/publication/262277120_Multiple_constraint_satisfaction_problems_using_the_A-star_A_search_algorithm_classroom_scheduling_with_preferences/links/5e57f530299bf1bdb8408b03/Multiple-constraint-satisfaction-problems-using-the-A-star-A-search-algorithm-classroom-scheduling-with-preferences.pdf}{Multiple constraint satisfaction problems using the A-star search algorithm}
    \bibitem{}
    \href{https://d1wqtxts1xzle7.cloudfront.net/68245993/14443-libre.pdf?1626958962=&response-content-disposition=inline%3B+filename%3DCourses_timetabling_based_on_hill_climbi.pdf&Expires=1714507660&Signature=JDd-gaDdE4E7s5P66Op535FyEuysvPoA5o4fcwak4NWT1sQNtVh50ooYa9eSrHu~aHR-Ulabenz910uDjozKpgHivCFkhipTT5bJJL9MensZTSib8TBVzGtBU9lycqCVwilt21Udi0DMgfhERDSgCcR4vO8hmlFlxOLg1exD1fNLE8kxt4S0H2Qfb3nQEHu3kLsicuJNk8wfF~eWpEDUlWsjW9J9IpxuLWP0OICQ7ZReWtx37dHnPN2PTVmFs815NMNi-jyLV0gWWBTAOL7Hn7A5VSg-l~s2e8RAoCjPAYQDRnnznIzOElRy2nTsgkW3mZDzlsntJ3lFur2KquwqNg__&Key-Pair-Id=APKAJLOHF5GGSLRBV4ZA}{Courses timetabling based on hill climbing algorithm}

    \end{thebibliography}
\end{document}